---
title: "SQMF - Summative Assessment"
author: "EXAM NUMBER"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
```

# Part I

Part I consists of a set of semi-guided exercises that assess your understanding of statistical concepts and R code.

There are a total of 10 exercises.

## Excercise 1

You are given three data file and you have to create one plot for each. Each data file requires you to read in the file, filter/transform the data and create a plot that shows a particular aspect of the data, as follows:

1. `data_e1_1.csv`: Plot centred speech rate against f0 at vowel mid-point, by condition and group, faceting by vowel.
2. `data_e1_2.csv`: Plot logged reaction times by language, environment, and age.
3. `data_e1_3.csv`: Plot proportions of incorrect vs correct responses across trial number in the easy and difficult condition, faceting by priming setting.

```{r e1-1}

```

```{r e1-2}

```

```{r e1-3}

```


# Exercise 2

The following plots are not appropriate for the type of data they show.

Briefly describe what is wrong with the plot and write code to create a more appropriate plot (looking at the data frame might help you).

```{r ex-2-1}

```

```{r ex-2-1-your-code}

```



```{r ex-2-2}

```

```{r ex-2-2-your-code}

```


## Exercise 3

Read the `ex_3.csv` file in R and obtain summary measures (central tendency: mean, median, or mode; and dispersion: standard deviation or min/max) for each variable in the data. Pick the appropriate measure of central tendency and dispersion for each variable.

## Exercise 4

Look at the following list of variables:

- Vowel duration.
- Formant values.
- Accuracy.
- Yes/no.
- Reaction times.
- Number of relative clauses.
- Scots vs English.

For each of these variables, specify whether it's (in principle) distributed according to a Gaussian, Bernoulli/Binomial or Poisson distribution.
To to do, create a *vector of strings*, where `"G"` = Gaussian, `"B"` = Bernoulli/Binomial, `"P"` = Poisson. Make sure you list the strings in the order in which the variables appear in the list above.

```{r ex-4}

```

**Extra question**: Are the Gaussian variables in the list exactly Gaussian? Briefly comment on this and if need be propose alternative distribution families for those variables and/or transformations one can apply to the variables to make them more Gaussian.

## Exercise 5

Look at the following table that represents the coding of a categorical predictor (assume we use the default alphabetical order). Explain what is wrong with it and write a new table with your solution (you can use <https://tablesgenerator.com/markdown_tables#> to format the markdown table).

|             | Bangladeshi | Mandarin | English |
|-------------|-------------|----------|---------|
| Bangladeshi | 1           | 0        | 0       |
| Mandarin    | 0           | 2        | 0       |
| English     | 0           | 0        | 3       |

## Exercise 6

Explain what are the benefits of centring continuous predictors and using sum coding with categorical predictors.

## Exercise 7

Read the `data_e7.csv` file. It contains the following columns:

- `subject`: the subject's ID.
- `response`: whether the subject has pressed the `Left` or `Right` button.
- `condition`: 
- 

Run a linear model that helps you answer the following question: Does ...?

## Exercise 8

Criticise the following linear model and run a more appropriate linear model to answer the following research question:

## Exercise 9

For each of the following 15 statements, say whether the statement is true or false.
To do so, create a *logical vector* with a list of `TRUE` and `FALSE`.






# Part II

Part II requires you to conduct a full data analysis of a data set, from reading data to transforming and plotting it, to linear modelling and reporting.

