---
title: "SQMF - Summative Assessment"
author: "YOUR EXAM NUMBER"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
```

# Overview

**IMPORTANT**: Remember to include your exam number as the author in the document preamble.

# Part I

Part I consists of a set of semi-guided exercises that assess your understanding of statistical concepts and R code.

There are a total of 10 exercises.

## Excercise 1

You are given three data file and you have to create one plot for each. Each data file requires you to read in the file, filter/transform the data and create a plot that shows a particular aspect of the data, as follows:

1. `data_e1_1.csv`: Plot centred speech rate against f0 at vowel mid-point, by condition and group, faceting by vowel.
2. `data_e1_2.csv`: Plot logged reaction times by language, environment, and age.
3. `data_e1_3.csv`: Plot proportions of incorrect vs correct responses across trial number in the easy and difficult condition, faceting by priming setting.

```{r e1-1}

```

```{r e1-2}

```

```{r e1-3}

```


# Exercise 2

The following plots are not appropriate for the type of data they show.

Briefly describe what is wrong with the plot and write code to create a more appropriate plot (looking at the data frame might help you).

```{r ex-2-1}

```

```{r ex-2-1-your-code}

```



```{r ex-2-2}

```

```{r ex-2-2-your-code}

```


## Exercise 3

Read the `ex_3.csv` file in R and obtain summary measures (central tendency: mean, median, or mode; and dispersion: standard deviation or min/max) for each variable in the data. Pick the appropriate measure of central tendency and dispersion for each variable.

## Exercise 4

Look at the following list of variables:

- Vowel duration.
- Formant values.
- Accuracy.
- Yes/no.
- Reaction times.
- Number of relative clauses.
- Scots vs English.

For each of these variables, specify whether it's (in principle) distributed according to a Gaussian, Bernoulli/Binomial or Poisson distribution.
To to do, create a *vector of strings*, where `"G"` = Gaussian, `"B"` = Bernoulli/Binomial, `"P"` = Poisson. Make sure you list the strings in the order in which the variables appear in the list above.

```{r ex-4}

```

**Extra question**: Are the Gaussian variables in the list exactly Gaussian? Briefly comment on this and if need be propose alternative distribution families for those variables and/or transformations one can apply to the variables to make them more Gaussian.

## Exercise 5

Look at the following table that represents the coding of a categorical predictor (assume we use the default alphabetical order). Explain what is wrong with it and write a new table with your solution (you can use <https://tablesgenerator.com/markdown_tables#> to format the markdown table).

|             | Bangladeshi | Mandarin | English |
|-------------|-------------|----------|---------|
| Bangladeshi | 1           | 0        | 0       |
| Mandarin    | 0           | 2        | 0       |
| English     | 0           | 0        | 3       |

## Exercise 6

Explain what are the benefits of centring continuous predictors and using sum coding with categorical predictors.

## Exercise 7

Read the `data_e7.csv` file. It contains the following columns:

- `subject`: the subject's ID.
- `response`: whether the subject has pressed the `Left` or `Right` button.
- `condition`: 
- 

Run a linear model that helps you answer the following question: Does ...?

## Exercise 8

Imagine you are asked to review a paper.
Below you can find the description of a mock study, including the details of the linear model the researchers have run. The results are not included.

> We recorded 50 subjects while they read 100 sentences on a screen. For each subject, half of the sentences were presented together with pictures of natural landscapes. The other half was presented together with pictures of urban landscapes. Background sounds were delivered via headphones to the subject in each trial of the natural and urban setting condition. For each setting (natural vs urban), half of the trials had natural sounds (birds, waterfalls, wind, waves, thunders) and half had urban sounds (traffic noise, sirens, people walking).

> For each trial we measured speech rate as number of syllables per second (syl/s).
The hypothesis is that speech rate will be faster in the urban setting trials relative to the natural setting trials. Moreover, the effect of background noise (natural vs urban sounds) will decrease speech rate in the natural setting but not in the urban setting. In other words, we expect the visual setting (natural vs urban) to prime speakers to slow down their speech rate, but we expect the auditory setting (natural vs urban) to make a difference only in the visual natural setting.

> To assess these expectations, we ran a linear model using a Gaussian distribution with visual setting (natural vs urban) as the outcome variable. We included the following predictors: speech rate (centred) and auditory setting (natural vs urban). In R syntax: `lm(visual ~ speech_rate_c + auditory)`.

Now criticise the analysis (i.e. explain what is wrong with it) and run a more appropriate linear model to assess the research hypotheses of the study based on the provided data (`data_e8.csv`).
Report the model specification and results of your linear model.

## Exercise 9

For each of the following 15 statements, say whether the statement is true or false.
To do so, create a *logical vector* with a list of `TRUE` and `FALSE`.






# Part II

Part II requires you to conduct a full data analysis of a data set, from reading data to transforming and plotting it, to linear modelling and reporting.

