---
title: "SQMF - Summative Assessment"
author: "EXAM NUMBER"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
```

# Part I

Part I consists of a set of semi-guided exercises that assess your understanding of statistical concepts and R code.

There are a total of 10 exercises.

## Excercise 1

You are given three data file and you have to create one plot for each. Each data file requires you to read in the file, filter/transform the data and create a plot that shows a particular aspect of the data, as follows:

1. `data_e1_1.csv`: Plot centred speech rate against f0 at vowel mid-point, by condition and group, faceting by vowel.
2. `data_e1_2.csv`: Plot logged reaction times by language, environment, and age.
3. `data_e1_3.csv`: Plot proportions of incorrect vs correct responses across trial number in the easy and difficult condition, faceting by priming setting.

```{r e1-1}

```

```{r e1-2}

```

```{r e1-3}

```


# Exercise 2

The following plots are not appropriate for the type of data they show.

Briefly describe what is wrong with the plot and write code to create a more appropriate plot (looking at the data frame might help you).

```{r ex-2-1}

```

```{r ex-2-1-your-code}

```



```{r ex-2-2}

```

```{r ex-2-2-your-code}

```


## Exercise 3

Read the `ex_3.csv` file in R and obtain summary measures (central tendency: mean, median, or mode; and dispersion: standard deviation or min/max) for each variable in the data. Pick the appropriate measure of central tendency and dispersion for each variable.

# Part II

Part II requires you to conduct a full data analysis of a data set, from reading data to transforming and plotting it, to linear modelling and reporting.

