---
title: "SQMF - Summative Assessment"
author: "YOUR EXAM NUMBER"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
```

# Instructions

**PLEASE READ CAREFULLY**

**IMPORTANT**: Remember to include your exam number as the author in the document preamble above.

**DUE THURSDAY 8 DECEMBER AT NOON**

**You must submit both to GitHub and to Learn**. No exceptions will be made.

The assessment is made of two parts:

- **Part I** is a series of semi-guided exercises on topics and procedures covered in class.
- **Part II** requires you to run a full data analysis of a data set from a set of four. You will have to pick one of the four data sets.

Complete each exercise by completing tasks, answering questions and/or providing code if required.
You will have to download the data for this assessment and copy it into the `data/` folder.
A link to download the data will be provided via an announcement.

When you are ready to submit:

1. Render the Rmd file to HTML.
2. Upload the Rmd and HTML file and the data files to GitHub. They should be uploaded in the correct folder (`code/` and `data/`).
3. Create a confirmation file to upload  to Learn. **The file must be named with your exam number only**: for example, `B231789.txt`, `B098273.docx`. **Failure to upload the file or not naming the file appropriately will correspond to a fail.**




# Part I

Part I consists of a set of semi-guided exercises that assess your understanding of statistical concepts and R code.

There are a total of 9 exercises.

## Excercise 1

You are given three data files and you have to create one plot for each. Each data file requires you to read in the file, filter/transform the data and create a plot that shows a particular aspect of the data, as follows:

1. `data_e1_1.csv`: Plot centred speech rate against f0 at vowel mid-point, by condition and group, faceting by vowel.
2. `data_e1_2.csv`: Plot logged reaction times by language, environment, and age.
3. `data_e1_3.csv`: Plot proportions of incorrect vs correct responses across trial number in the easy and difficult condition, faceting by priming setting.

```{r e1-1}

```

```{r e1-2}

```

```{r e1-3}

```


## Exercise 2

The following plots are not appropriate for the type of data they show.

Briefly describe what is wrong with the plot and write code to create a more appropriate plot (looking at the data frame might help you).

```{r e2-1}
data_e2_1 %>%
  ggplot(aes(reponse, condition)) +
  geom_point()
```

```{r e2-1-your-code}

```



```{r e2-2}
data_e2_2 %>%
  ggplot(aes(vot, fill = voicing)) +
  geom_bar()
```

```{r e2-2-your-code}

```


## Exercise 3

Read the `ex_3.csv` file in R and obtain summary measures (central tendency: mean, median, or mode; and dispersion: standard deviation or min/max) for each variable in the data. Pick the appropriate measure of central tendency and dispersion for each variable.

## Exercise 4

Look at the following list of variables:

- Vowel duration.
- Formant values.
- Accuracy.
- Yes/no.
- Reaction times.
- Number of relative clauses.
- Scots vs English.

For each of these variables, specify whether it's (in principle) distributed according to a Gaussian, Bernoulli/Binomial or Poisson distribution.
To to do, create a *vector of strings*, where `"G"` = Gaussian, `"B"` = Bernoulli/Binomial, `"P"` = Poisson. Make sure you list the strings in the order in which the variables appear in the list above.

```{r e4}

```

## Exercise 5

Look at the following table that represents the coding of a categorical predictor (assume we use the default alphabetical order). Explain what is wrong with it and write a new table with your solution (you can use <https://tablesgenerator.com/markdown_tables#> to format the markdown table).

|             | Bangladeshi | Mandarin | English |
|-------------|-------------|----------|---------|
| Bangladeshi | 1           | 0        | 0       |
| Mandarin    | 0           | 2        | 0       |
| English     | 0           | 0        | 3       |

## Exercise 6

Explain what are the benefits of centring continuous predictors and using sum coding with categorical predictors.

## Exercise 7

Imagine you run the following study:

Participants are asked to listen to nonce words and choose whether they think the word refers to a small or a big object.
Half of the words are of the `kiki` type (back consonants and high front vowels), while half are of the `baba` type (front consonant and low vowels).
The expectation is that `kiki` words should elicit more `small` responses and `baba` words more `big` responses.
We also recorder reaction time and we expect shorter reaction times to correlate with a greater effect of hearing a `kiki` word vs a `baba` word.

Read the `data_e7.csv` file. It contains the following columns:

- `subject`: the subject's ID.
- `response`: whether the subject has chosen `small` or `big`.
- `condition`: `kiki` vs `baba` word.
- `RT`: reaction time in ms.

Make sure to change columns to factors if needed and to specify the order of levels.

```{r e7-1}

```


Run a linear model that helps you answer the following questions:

1. Do `baba` words elicit more `big` responses than `kiki` words?
2. Is the effect of `baba` words on response greater with shorter reaction times?

```{r e7-2}

```

Report the model specification and the results. Also include a plot of the model results.

```{r e7-3}

```


## Exercise 8

Imagine you are asked to review a paper.
Below you can find the description of a mock study, including the details of the linear model the researchers have run. The results are not included.

> We recorded 50 subjects while they read 100 sentences on a screen. For each subject, half of the sentences were presented together with pictures of natural landscapes. The other half was presented together with pictures of urban landscapes. Background sounds were delivered via headphones to the subject in each trial of the natural and urban setting condition. For each setting (natural vs urban), half of the trials had natural sounds (birds, waterfalls, wind, waves, thunders) and half had urban sounds (traffic noise, sirens, people walking).

> For each trial we measured speech rate as number of syllables per second (syl/s).
The hypothesis is that speech rate will be faster in the urban setting trials relative to the natural setting trials. Moreover, the effect of background noise (natural vs urban sounds) will decrease speech rate in the natural setting but not in the urban setting. In other words, we expect the visual setting (natural vs urban) to prime speakers to slow down their speech rate, but we expect the auditory setting (natural vs urban) to make a difference only in the visual natural setting.

> To assess these expectations, we ran a linear model using a Gaussian distribution with visual setting (natural vs urban) as the outcome variable. We included the following predictors: speech rate (centred) and auditory setting (natural vs urban). In R syntax: `lm(visual ~ speech_rate_c + auditory)`.

Now criticise the analysis (i.e. explain what is wrong with it) and run a more appropriate linear model to assess the research hypotheses of the study based on the provided data (`data_e8.csv`).
Report the model specification and results of your linear model.

## Exercise 9

For each of the following 15 statements, say whether the statement is true or false.
To do so, create a *logical vector* with a list of `TRUE` and `FALSE`.


1. A p-value is the probability of the null hypothesis being true.

2. Sum-coding a predictor sets the intercept to the grand mean for that predictor.

3. A significant result is evidence for the alternative hypothesis.

4. A 95% Confidence Interval indicates the probability of having obtained the true mean.

5. The most appropriate distribution family for a dichotomous outcome variable is the Bernoulli family.

6. The ultimate goal of statistics is to assess significance using p-values.

7. Strip charts are appropriate to visualise the individual observations of a continuous variable.

8. Variables included as random effects in a linear model cannot also be included as fixed effects.

9. Linear models assume that the observations in the data are independent.

10. The Student's *t*-statistic is a standardised measure for the difference between two means.

11. About 68% of the data in a Gaussian distribution is contained within the range marked by mean - 1 SD and mean + 1 SD.

12. The alpha level can only be 0.05.

13. Reproducibility and replicability are the same thing.

14. A study is *robust* when the same results are obtained with the same data but a different analysis.

15. The types of predictor variables in a linear model decide which distribution family to use in the model.




# Part II

Part II requires you to conduct a full data analysis of a data set, from reading data to transforming and plotting it, to linear modelling and reporting.

